---
title: "Heatmap plotting"
output: html_notebook
---
```{r}
library(tidyverse)
library(scales)
library(egg)
```
Set `data_path` to the directory where you have the API tables and party rating CSV files. This assumes file names have not been modified, in case they have been, adjust accordingly. Set `plot_path` to where you want plots be saved.

```{r}
data_path = ''
plot_path = ''
```

## API plot

Reads API index values and other values from a table and outputs a plot with API values and other values by year.

```{r}
indices <- read.csv(paste(data_path, 'API_tables2.csv', sep = ''))
```

```{r}
to_plot <- indices %>%
  select(-X) %>%
  rename('2007' = 'X2007', '2011' = 'X2011', '2015' = 'X2015', '2019' = 'X2019') %>%
  pivot_wider(everything(), names_from = 'point', values_from = '2007':'2019') %>%
  pivot_longer(starts_with('2'), names_to = c('year', '.value'), names_pattern = '(.*)_(.*)') %>%
  #filter(attribute == 'API') %>%
  mutate(year = as.numeric(year))
```

```{r}
first_plot <- to_plot %>% filter(!attribute %in% c('API', 'weighted_API', 'raw_polarization', 'mean AP'))

p1 <- ggplot(data = first_plot, aes(x = year, group = attribute, fill = attribute)) +
  geom_ribbon(aes(x = year, ymin = low, ymax = up),
              alpha = 0.3) +
  geom_line(aes(x = year, y = mean, color = attribute)) +
  #theme(legend.position = 'bottom') +
  #scale_fill_brewer(palette = "Set2") +
  scale_fill_brewer(palette = "Set2", labels = c('Inparty mean',
                                 'Outparty maximum',
                                 'Outparty mean',
                                 'Outparty minimum',
                                 'Party mean')) +
  scale_color_brewer(palette = "Set2", labels = c('Inparty mean',
                                 'Outparty maximum',
                                 'Outparty mean',
                                 'Outparty minimum',
                                 'Party mean')) +
  theme(legend.title = element_blank()) +
  ylab('Rating') +
  xlab('Year') +
  ylim(c(0,10)) +
  scale_x_continuous(breaks = c(2007, 2011, 2015, 2019), limits = c(2007, 2019))

second_plot <- to_plot %>% filter(attribute == 'API') 

p2 <- ggplot(data = second_plot, aes(x = year, group = attribute)) +
  geom_ribbon(data = second_plot, aes(x = year, ymin = low, ymax = up, fill = attribute),
              alpha = 0.3) +
  geom_line(aes(x = year, y = mean, group = attribute, color = attribute)) +
  theme(legend.position = 'none') +
  ylab('API') +
  xlab('Year') +
  scale_x_continuous(breaks = c(2007, 2011, 2015, 2019), limits = c(2007, 2019))
```

```{r}
ggsave(paste(plot_path, 'indices.png', sep = ''),
       ggarrange(p2, p1, ncol = 2), width = 7, height = 3)
ggsave(paste(plot_path, 'indices.tiff', sep = ''),
       ggarrange(p2, p1, ncol = 2), width = 7, height = 3, dpi = 300)
```

## Opinion heatmaps

```{r}
ops_2007 <- read.csv(paste(data_path, 'opinions_2007.csv', sep = ""))
ops_2011 <- read.csv(paste(data_path, 'opinions_2011.csv', sep = ""))
ops_2015 <- read.csv(paste(data_path, 'opinions_2015.csv', sep = ""))
ops_2019 <- read.csv(paste(data_path, 'opinions_2019.csv', sep = ""))
```

```{r}
m1 <- ops_2007 %>%
  pivot_longer(CD:SPP, names_to = 'party', values_to = "opinion") %>%
  rename('partisan_group' = 'X') %>%
  mutate_at(vars(party), funs(factor(., levels = c("CD", "CPF", "NCP", "FP", "SPP", "SDP", "LA", "GL")))) %>%
  mutate_at(vars(partisan_group), funs(factor(., levels = c("CD", "CPF", "NCP", "FP", "SPP", "SDP", "LA", "GL"))))
m1['year'] = 2007

m2 <- ops_2011 %>%
  pivot_longer(CD:SPP, names_to = 'party', values_to = "opinion") %>%
  rename('partisan_group' = 'X') %>%
  mutate_at(vars(party), funs(factor(., levels = c("CD", "CPF", "NCP", "FP", "SPP", "SDP", "LA", "GL")))) %>%
  mutate_at(vars(partisan_group), funs(factor(., levels = c("CD", "CPF", "NCP", "FP", "SPP", "SDP", "LA", "GL"))))
m2['year'] = 2011

m3 <- ops_2015 %>%
  pivot_longer(CD:SPP, names_to = 'party', values_to = "opinion") %>%
  rename('partisan_group' = 'X') %>%
  mutate_at(vars(party), funs(factor(., levels = c("CD", "CPF", "NCP", "FP", "SPP", "SDP", "LA", "GL")))) %>%
  mutate_at(vars(partisan_group), funs(factor(., levels = c("CD", "CPF", "NCP", "FP", "SPP", "SDP", "LA", "GL"))))
m3['year'] = 2015

m4 <- ops_2019 %>%
  pivot_longer(CD:SPP, names_to = 'party', values_to = "opinion") %>%
  rename('partisan_group' = 'X') %>%
  mutate_at(vars(party), funs(factor(., levels = c("CD", "CPF", "NCP", "FP", "SPP", "SDP", "LA", "GL")))) %>%
  mutate_at(vars(partisan_group), funs(factor(., levels = c("CD", "CPF", "NCP", "FP", "SPP", "SDP", "LA", "GL"))))
m4['year'] = 2019

m_combi <- rbind(m1, m2, m3, m4)
```

```{r}
opinion_heatmap <- m_combi %>%
  ggplot(aes(party, partisan_group, fill = opinion)) +
    geom_tile() +
    scale_fill_distiller(name = 'Rating', palette = 'RdBu',
                         limits = c(0, 10), breaks = c(0, 2.5, 5, 7.5, 10), labels = c('0 (Cold)', '2.5', '5', '7.5', '10 (Warm)')) +
    geom_text(aes(label = round(opinion, 1))) +
    theme_minimal() + 
    facet_wrap(~ year) +
    ylab('Partisan group') +
    xlab('Party')

ggsave(paste(plot_path, 'opinion_heatmap.png', sep = ''), width = 7, height = 4, opinion_heatmap)
ggsave(paste(plot_path, 'opinion_heatmap.tiff', sep = ''), width = 7, height = 4, dpi = 300, opinion_heatmap)
```

## Distance heatmaps

```{r}
dists_2007 <- read.csv(paste(data_path, 'distances_2007.csv', sep = ""))
dists_2011 <- read.csv(paste(data_path, 'distances_2011.csv', sep = ""))
dists_2015 <- read.csv(paste(data_path, 'distances_2015.csv', sep = ""))
dists_2019 <- read.csv(paste(data_path, 'distances_2019.csv', sep = ""))
```

```{r}
dm1 <- dists_2007 %>%
  pivot_longer(CD:SPP, names_to = 'party', values_to = "opinion") %>%
  rename('partisan_group' = 'X') %>%
  mutate_at(vars(party), funs(factor(., levels = c("CD", "CPF", "NCP", "FP", "SPP", "SDP", "LA", "GL")))) %>%
  mutate_at(vars(partisan_group), funs(factor(., levels = c("CD", "CPF", "NCP", "FP", "SPP", "SDP", "LA", "GL"))))
dm1['year'] = 2007

dm2 <- dists_2011 %>%
  pivot_longer(CD:SPP, names_to = 'party', values_to = "opinion") %>%
  rename('partisan_group' = 'X') %>%
  mutate_at(vars(party), funs(factor(., levels = c("CD", "CPF", "NCP", "FP", "SPP", "SDP", "LA", "GL")))) %>%
  mutate_at(vars(partisan_group), funs(factor(., levels = c("CD", "CPF", "NCP", "FP", "SPP", "SDP", "LA", "GL"))))
dm2['year'] = 2011

dm3 <- dists_2015 %>%
  pivot_longer(CD:SPP, names_to = 'party', values_to = "opinion") %>%
  rename('partisan_group' = 'X') %>%
  mutate_at(vars(party), funs(factor(., levels = c("CD", "CPF", "NCP", "FP", "SPP", "SDP", "LA", "GL")))) %>%
  mutate_at(vars(partisan_group), funs(factor(., levels = c("CD", "CPF", "NCP", "FP", "SPP", "SDP", "LA", "GL"))))
dm3['year'] = 2015

dm4 <- dists_2019 %>%
  pivot_longer(CD:SPP, names_to = 'party', values_to = "opinion") %>%
  rename('partisan_group' = 'X') %>%
  mutate_at(vars(party), funs(factor(., levels = c("CD", "CPF", "NCP", "FP", "SPP", "SDP", "LA", "GL")))) %>%
  mutate_at(vars(partisan_group), funs(factor(., levels = c("CD", "CPF", "NCP", "FP", "SPP", "SDP", "LA", "GL"))))
dm4['year'] = 2019

dm_combi <- rbind(dm1, dm2, dm3, dm4)
```

```{r}
distance_heatmap <- dm_combi %>%
  ggplot(aes(party, partisan_group, fill = opinion)) +
    geom_tile() +
    scale_fill_distiller(name = 'Distance', palette = 'RdBu',
                         direction = 1, limits = c(0, 10), breaks = c(10, 7.5, 5, 2.5, 0), labels = c('10 (Cold)', '7.5', '5', '2.5', '0 (Warm)')) +
    geom_text(aes(label = round(opinion, 1))) +
    theme_minimal() + 
    facet_wrap(~ year) +
    ylab('Partisan group') +
    xlab('Party')

ggsave(paste(plot_path, 'distance_heatmap.png', sep = ''), width = 7, height = 4, distance_heatmap)
ggsave(paste(plot_path, 'distance_heatmap.tiff', sep = ''), width = 7, height = 4, dpi = 300, distance_heatmap)
```